Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2022-11-21 analysis/paper/paper.qmd
Render qmd file
minor edits to some figures

eco_box <- ggplot(points_in_veg_zones_all) +
  aes(ecoregion,
      min_tcsa) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.3,
                   size = 2.3) +
  xlab("Vegetation") + 
  ylab("TCSA")

@@

for_box <- ggplot(points_in_veg_zones_forest_box) +
  aes(Zone,
      min_tcsa) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.3,
                   size = 2.3) +
  xlab("Ecoregion") + 
  ylab("TCSA")

plot_grid(eco_plot, for_plot, eco_box, for_box, labels = c('A', 'C', 'B', 'D'))
```

@@

#vegetation region#
points_in_veg_zones_all$ecoregion = as.factor(points_in_veg_zones_all$ecoregion)
eco_aov <- aov(min_tcsa ~ ecoregion, points_in_veg_zones_all)
summary(eco_aov)
eco_tur<- TukeyHSD(eco_aov)
# p-value
# simple version
# eco_lgm <- round(summary(eco_aov)[[1]][[5]][1], digits = 3)

eco_lgm_test <- Anova(lm(min_tcsa ~  ecoregion, points_in_veg_zones_all))
# describe.Anova(eco_lgm_test, 1, 2)
# plot(eco_tur)
# plot with ggplot
eco_tur_df <- as.data.frame(eco_tur$ecoregion)

eco_tur_box <- ggplot(
  data = eco_tur_df,
  mapping = aes(
    x = row.names(eco_tur_df),
    lower = lwr,
    middle = diff,
    upper = upr,
    ymin = lwr,
    ymax = upr
    )
  ) +
  geom_boxplot(stat = "identity")+
  xlab("Vegetation zones")+
  theme(axis.text.x = element_text(angle = 45, size = 6)) +
  theme(axis.text.y = element_text(size = 6))

#eco region#
for_aov <- aov(min_tcsa ~ as.factor(Zone), points_in_veg_zones_forest_box)
summary(for_aov)
for_tur<- TukeyHSD(for_aov)
# p-value
# simple version
# for_lgm <- round(summary(for_aov)[[1]][[5]][1], digits = 3)
for_lgm_test <- Anova(lm(min_tcsa ~ as.factor(Zone), points_in_veg_zones_forest_box))
# describe.Anova(for_lgm_test, 1, 2)
# plot(for_tur)

for_tur_df <- as.data.frame(for_tur$`as.factor(Zone)`)

for_tur_box <- ggplot(
  data = for_tur_df,
  mapping = aes(
    x = row.names(for_tur_df),
    lower = lwr,
    middle = diff,
    upper = upr,
    ymin = lwr,
    ymax = upr
    )
  ) +
  geom_boxplot(stat = "identity") +
  xlab("Eco-region zones") +
  theme(axis.text.x = element_text(angle = 45, size = 6)) +
  theme(axis.text.y = element_text(size = 6))

# ggplot(for_tur_df) +
#   geom_errorbar(aes(
#     x = row.names(for_tur_df),
#     ymin = lwr,
#     ymax = upr
#     ))

plot_grid(eco_tur_box, for_tur_box, labels = c('A', 'B'))

@@

In order to explore spatial patterns as well as the relationship between the function of the tools and the environment, we located individual assemblages on maps that depict eco-regional boundaries. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) of vegetation map or Central inland (CI) (@fig-gis-tcsa-sites) of eco-region map. We observed that TCSA values are strongely associated with the zones (Vegetation zones: `apastats::describe.Anova(eco_lgm_test, 1, 2)`, Ecoregion zones: `apastats::describe.Anova(for_lgm_test, 1, 2)`). To explore the TCSA distribution within the eco-regional zones, we excluded zones in the ecoregion map that have no stemmed points (@fig-gis-tcsa-sites panel D). We found no clear functional pattern throughout the zones, showing a wide variation in TCSA regardless of the density of stemmed point in each zone (@fig-gis-tcsa-sites panel B and D).

@@

Marlize Lombard: Conceptualization, Methodology, Writing - Review & Editing, Supervision

Donghee Chong: Data curation, Review & Editing

>>>>

eco_box <- ggplot(points_in_veg_zones_all) +
  aes(ecoregion,
      min_tcsa) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.3,
                   size = 2.3) +
  xlab("Vegetation") + 
  ylab("TCSA") +
  theme_minimal()

@@

for_box <- ggplot(points_in_veg_zones_forest_box) +
  aes(Zone,
      min_tcsa) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.3,
                   size = 2.3) +
  xlab("Ecoregion") + 
  ylab("TCSA") +
  theme_minimal()

@@

#vegetation region#
points_in_veg_zones_all$ecoregion = as.factor(points_in_veg_zones_all$ecoregion)
eco_aov <- aov(min_tcsa ~ ecoregion, points_in_veg_zones_all)

eco_lgm_test <- Anova(lm(min_tcsa ~ ecoregion, points_in_veg_zones_all))

eco_lgm_test_pretty <- apastats::describe.Anova(eco_lgm_test, 1, 2)

eco_tur <- TukeyHSD(eco_aov)

# plot Tukey HSD
eco_tur_box <- 
eco_tur$ecoregion %>% 
  as_tibble(rownames = "zones") %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones)) +
  geom_point(aes(diff,
                 zones),
             size = 4) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\nvegetation zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8)

#eco region#
for_aov <- aov(min_tcsa ~ as.factor(Zone), points_in_veg_zones_forest_box)
# summary(for_aov)
for_lgm_test <- Anova(lm(min_tcsa ~ as.factor(Zone), points_in_veg_zones_forest_box))

for_lgm_test_pretty <- apastats::describe.Anova(for_lgm_test, 1, 2)

for_tur <- TukeyHSD(for_aov)

# plot Tukey HSD
for_tur_box <- 
for_tur$`as.factor(Zone)` %>% 
  as_tibble(rownames = "zones") %>% 
  mutate(zones = fct_reorder(zones, diff)) %>% 
  ggplot() +
  geom_linerange(aes(xmin = lwr, 
                     xmax = upr,
                     y = zones)) +
  geom_point(aes(diff,
                 zones),
             size = 4) +
  geom_vline(xintercept = 0, colour = "red") +
  ylab("Pairwise comparison\nof TSCA in different\necological zones") +
  xlab("Differences in means") +
  theme_minimal(base_size = 8)

plot_grid(eco_plot, 
          for_plot, 
          eco_box, 
          for_box, 
          eco_tur_box, 
          for_tur_box,
          labels = c('A', 'C', 'B', 'D', 'E', 'F'),
          ncol = 2)

@@

In order to explore spatial patterns as well as the relationship between the function of the tools and the environment, we located individual assemblages on maps that depict eco-regional boundaries. Our results show that stemmed points are only located in certain zones, Central Temperature Zone (CT) of vegetation map or Central inland (CI) (@fig-gis-tcsa-sites) of eco-region map. We observed that TCSA values are strongely associated with the zones (Vegetation zones: `r eco_lgm_test_pretty`, Ecoregion zones: `r for_lgm_test_pretty`). To explore the TCSA distribution within the eco-regional zones, we excluded zones in the ecoregion map that have no stemmed points (@fig-gis-tcsa-sites panel D). We found no clear functional pattern throughout the zones, showing a wide variation in TCSA regardless of the density of stemmed point in each zone (@fig-gis-tcsa-sites panel B and D).

@@

Marlize Lombard: Conceptualization, Methodology, Supervision

Donghee Chong: Data curation
